Reanalysis Step 4: Outcome Analysis with Three-Way Interaction

### 2025-08-17 reanalysis-4-outcome_analysis.qmd
twang

# Test 1: Joint test of three-way interaction
threeway_terms_test <- grep("ever_lapse.*:dr_severity.*:any_treatment", 
                            names(coef(model_for_tests)), value = TRUE)

@@

  joint_test_3way <- linearHypothesis(
    model_for_tests,
    hypothesis_3way,
    vcov = vcov(model_for_tests)
  )

  message("\n1. Three-way interaction test (H0: all three-way terms = 0):")
  message("   Chi-sq = ", round(joint_test_3way$Chisq[2], 3))
  message("   df = ", joint_test_3way$Df[2])
  message("   p-value = ", format.pval(joint_test_3way$`Pr(>Chisq)`[2], digits = 4, eps = 0.0001))

  if(joint_test_3way$`Pr(>Chisq)`[2] < 0.05) {
    message("   Interpretation: The effect of lapsing varies by BOTH DR severity AND treatment status")
  } else {
    message("   Interpretation: No significant three-way interaction detected")
  }
}

# Test 2: Joint test of lapse × DR interaction (averaging over treatment)
lapse_dr_terms <- grep("ever_lapse.*:dr_severity(?!.*:any_treatment)", 
                      names(coef(model_for_tests)), value = TRUE, perl = TRUE)

@@

  joint_test_lapse_dr <- linearHypothesis(
    model_for_tests,
    hypothesis_lapse_dr,
    vcov = vcov(model_for_tests)
  )

  message("\n2. Lapse × DR severity interaction test:")
  message("   Chi-sq = ", round(joint_test_lapse_dr$Chisq[2], 3))
  message("   df = ", joint_test_lapse_dr$Df[2])
  message("   p-value = ", format.pval(joint_test_lapse_dr$`Pr(>Chisq)`[2], digits = 4, eps = 0.0001))
}

# Test 3: Joint test of lapse × treatment interaction (averaging over DR)
lapse_tx_terms <- grep("ever_lapse.*:any_treatment(?!.*:dr_severity)", 
                      names(coef(model_for_tests)), value = TRUE, perl = TRUE)

@@

  joint_test_lapse_tx <- linearHypothesis(
    model_for_tests,
    hypothesis_lapse_tx,
    vcov = vcov(model_for_tests)
  )

  message("\n3. Lapse × treatment interaction test:")
  message("   Chi-sq = ", round(joint_test_lapse_tx$Chisq[2], 3))
  message("   df = ", joint_test_lapse_tx$Df[2])
  message("   p-value = ", format.pval(joint_test_lapse_tx$`Pr(>Chisq)`[2], digits = 4, eps = 0.0001))
}
```

>>>>

# Use the first model to get term names (should be same across all imputations)
first_model <- models[[1]]

# Test 1: Joint test of three-way interaction
threeway_terms_test <- grep("ever_lapse.*:dr_severity.*:any_treatment", 
                            names(coef(first_model)), value = TRUE)

@@

  # Pool test across imputations
  pooled_test_3way <- pool_chi_square_tests(models, hypothesis_3way, "Three-way interaction")

  message("\n1. Three-way interaction test (H0: all three-way terms = 0):")
  message("   Chi-sq = ", round(pooled_test_3way$D_bar, 3))
  message("   df = ", pooled_test_3way$df)
  message("   p-value = ", format.pval(pooled_test_3way$p_value, digits = 4, eps = 0.0001))

  if(pooled_test_3way$p_value < 0.05) {
    message("   Interpretation: The effect of lapsing varies by BOTH DR severity AND treatment status")
  } else {
    message("   Interpretation: No significant three-way interaction detected")
  }
}

# Test 2: Joint test of lapse × DR interaction (averaging over treatment)
lapse_dr_terms <- grep("ever_lapse.*:dr_severity(?!.*:any_treatment)", 
                      names(coef(first_model)), value = TRUE, perl = TRUE)

@@

  # Pool test across imputations
  pooled_test_lapse_dr <- pool_chi_square_tests(models, hypothesis_lapse_dr, "Lapse × DR interaction")

  message("\n2. Lapse × DR severity interaction test:")
  message("   Chi-sq = ", round(pooled_test_lapse_dr$D_bar, 3))
  message("   df = ", pooled_test_lapse_dr$df)
  message("   p-value = ", format.pval(pooled_test_lapse_dr$p_value, digits = 4, eps = 0.0001))
}

# Test 3: Joint test of lapse × treatment interaction (averaging over DR)
lapse_tx_terms <- grep("ever_lapse.*:any_treatment(?!.*:dr_severity)", 
                      names(coef(first_model)), value = TRUE, perl = TRUE)

@@

  # Pool test across imputations
  pooled_test_lapse_tx <- pool_chi_square_tests(models, hypothesis_lapse_tx, "Lapse × treatment interaction")

  message("\n3. Lapse × treatment interaction test:")
  message("   Chi-sq = ", round(pooled_test_lapse_tx$D_bar, 3))
  message("   df = ", pooled_test_lapse_tx$df)
  message("   p-value = ", format.pval(pooled_test_lapse_tx$p_value, digits = 4, eps = 0.0001))
}
```
